---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


##Loading and preprocessing the data
```{r}
unzip ("activity.zip", exdir = "./")
dt <- read.table("activity.csv", sep =",", header = TRUE)
```

##What is mean total number of steps taken per day?
###Total number of steps taken per day
```{r}
mean_day <- aggregate(dt[,1], list(dt$date), sum)
```
###Histogram of the total number of steps taken each day
```{r}
hist(mean_day$x, main = "Histogram of the total number of steps taken each day", xlab="Steps per day",col="orange")
```

###Mean and median of the total number of steps taken per day
```{r}
mean(mean_day$x,na.rm=TRUE)
median(mean_day$x,na.rm=TRUE)
```

## What is the average daily activity pattern?
###A time series plot of the 5-minute interval and the average number of steps taken, averaged across all days
```{r}
intm <- aggregate(dt[,1], by=list(dt$interval), FUN=mean, na.rm=TRUE)
plot(intm$Group.1,intm$x,type="l", main="Time Series Plot of the intervals and the average number of steps", xlab="Intervals", ylab="Average number of steps",col="pink")
```

###The 5-minute interval, on average across all the days in the dataset, containing the maximum number of steps
```{r}
intm[which.max(intm$x),1]
```

## Imputing missing values
##Total number of missing values in the dataset
```{r}
length(which(is.na(dt)))
```
###Strategy for filling in all of the missing values in the dataset using the mean for that 5-minute interval and creating a new dataset "d" with all filled in values
```{r}
dtf <- read.csv("activity.csv")
library(plyr)
d <- mutate(dtf, steps = replace(dtf$steps, is.na(dtf$steps), intm$x))
```

###Histogram of the total number of steps taken each day
```{r}
d_f <- aggregate(d[,1], list(d$date), sum, rm.na=TRUE)
hist(d_f$x, main = "Histogram of the total number of steps taken each day", xlab="Steps per Day",col="brown")
```

###Mean and median of the total number of steps taken per day
```{r}
mean(d_f$x)
median(d_f$x)
```

###The mean has increased slightly and the median is now same as the mean.

## Are there differences in activity patterns between weekdays and weekends?

###New factor variable in the dataset with two levels weekday and weekend indicating whether a given date is a weekday or weekend day
```{r}
d$weekday <- weekdays(as.Date(d$date,"%Y-%m-%d"))
d$factor_week <- ifelse(grepl('^S', d$weekday), "weekend", "weekday")

d_weekday <- subset(d,d$factor_week=="weekday")
d_weekend <- subset(d,d$factor_week=="weekend")

intm_weekday <- aggregate(d_weekday[,1], by=list(d_weekday$interval), FUN=mean)
intm_weekend <- aggregate(d_weekend[,1], by=list(d_weekend$interval), FUN=mean)
```

###Time series plot of the differences in activities on the weekdays and the weekends
```{r}
par(mfrow=c(2,1),mar=c(4,4,2,2))
plot(intm_weekday$Group.1,intm_weekday$x,type="l", main="Weekdays",xlab="",ylab="",col="blue")
plot(intm_weekend$Group.1,intm_weekend$x,type="l", main="Weekends", xlab="Intervals", ylab="Average number of steps",col="red")
```
